Quantum-Classical Phase Transition for Exciton-Polariton Josephson Junctions

### 2017-02-27 writing/phase_transition/phase_transition.tex
pre-final version

\author{M. E. Lebedev}
\email{m.lebedev@lmc.ifmo.ru}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{D. A. Dolinina}
\email{dasha.doly@gmail.com}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{A. P. Alodjants}
\email{alexander\_ap@list.ru}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

@@

\begin{abstract}
We consider the quantum phase problem for low branch exciton polariton condensate Josephson junctions operating at finite temperatures and polariton lifetime. The process of quantum tunneling we characterized by thermon lifetime. We have shown that classical (thermal) activation and quantum tunneling for the phase  represent two regimes depending on polariton condensate parameters and distance between the junctions. The transition between this two regimes   exhibits universal features of first or second order phase transition depending on the potential shape for the relative phase. Notably, scale of the crytical temperature  characterized by polariton interaction strength only. We have shown that polariton condensate junction device runs in classical way for ten of Kelvins temperatures and typical polariton densities used in current experiments with narrow band semiconductor samples. Purely  quantum phase regime can be obtained even in the presence of dissipation at one or two order lower temperatures and relatively short operating time scales (few of picoseconds), and   opens kindly new possibilities for exciton polariton BEC applications in scalable quantum computing and quantum metrology.   
\end{abstract}

@@

For last decade investigations of exciton polariton Bose-Einstein condensates (BEC) in various type of semiconductor microstructures represent one of mainstreams of current photonic and semiconductor technologies, see e.g. \cite{Sanvitto,Guillet}.
The microcavity exciton polaritons are bosonic quasiparticles representing  admixtures of the quantized cavity mode and quantum wells excitons.
Practically, semiconductor microcavities are promising for various optoelectronic applications where quantum  matter-field interface plays crucial role.
In particular, demonstrations of polariton laser with electrical pump \cite{Bhattacharya,Schneider}, amplifiers \cite{Niemietz}, switches \cite{Amo_2010}, transistors \cite{Ballarini}, polariton circuits and optical logic elements \cite{Sturm,Liew}.
Although exciton polariton condensates exhibits Bose-Einstein statistics under the phase transition condition and macroscopic occupation of the ground state at certain pumping rate that is less than for convenient lasers, they are not in true thermodynamic equilibrium state \cite{Byrnes_2014,Sun}.

Non-equilibrium, or, thermodynamically quasi-equilibrium features of exciton polariton condensates play certain role in various manifestations of their collective (many body) quantum states such as superfluidity \cite{Carusotto_2013,Amo_2009}, quantized vortices \cite{Lagoudakis_2008,Lagoudakis_2009}, soliton formation \cite{Sich}, Josephson oscillations and macroscopic self-trapping \cite{Abbarchi,Lagoudakis_2010}.
Since exciton polariton BEC's in present experiments examined at high enough (up to the room) temperatures the quantum character of such a states should be examined even at equilibrium.

The problem of distinguishability of statistically classical and quantum regimes for exciton polariton condensates behavior is very actual for their practical purposes, see e.g. \cite{Dominici}; especially for their possible applications for quantum information science purposes.
Fast switching properties (the typical switching time of a few picoseconds), relatively strong nonlinear response and flexibility to external optical and/or electrical pump, spin degrees of freedom made microcavity polaritons potentially very promising for quantum computation and quantum information processing objectives, see, e.g. \cite{Demirchyan,Pagel,Kyriienko,Solnyshkov_2015, Dominici}.
Especially it is important to stress quantum annealing problem \cite{Yan} that is very actual now and where collective (bosonic) character of polariton BEC could be used.
In particular, as it is shown recently in \cite{Lewenstein} global minimum of energy landscape can be reached at finite temperatures via a combination of thermal annealing and quantum tunneling.

Another promising area of potential applications of quantum exciton polariton states is connected with quantum metrology purposes for which designing of quantum interferometers and gyroscopes possessing high precision measurements with quantum phase states is utilized, cf. \cite{Pezze, Gulevich}.

In the paper we consider generic problem  of quantum-classical phase transition (see e.g.\cite{Chudnovsky_1997}) with exciton polariton condensates occurring in semiconductor microcavity in the presence of Josephson effect, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010}.
In general this problem relates to dissipative tunneling problem or, tunneling at finite (non-zero) temperatures models, cf. \cite{Caldeira, Larkin, Riseborough}.
Traditionally  seminal studies on this topic are based on consideration of superconductor devices \cite{Ankerhold} where macroscopic tunneling plays essential role.
In particular, it is worth to mention here Schr\"odinger cat state formation \cite{Leggett} and design of quantum gates for quantum computing with superconductor qubits where macroscopic quantum coherent phase properties are used \cite{Makhlin}.
Latterly the problem of quantum tunneling and quantum-classical phase transition problem have been applied to other two-level (spin) systems, like atomic BEC's \cite{Zhang}, and small ferromagnetic particles, \cite{Owerre}.
Strictly speaking main discussion is focused on investigation of quantum-classical escape rate transition that takes place in the presence of potential barrier at finite temperatures.

In the paper we propose \textit{extended model} for exciton polariton Josephson junctions that relates to so-called non-standard Bose-Hubbard models (see \cite{Dutta} and references therein) when energy of polariton-polariton scattering contributes into the tunneling parameter, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010, Solnyshkov_2008, Sarchi}.
In particularly, we pay our attention to temperature dependent quantum critical phenomena occurring in the presence of macroscopic tunneling.
We aimed at elucidation of necessary criteria to obtain quantum phase regimes with trapped exciton polaritons suitable for creation of exciton polariton qubit gates.

This paper is arranged as follows.
In Sec. \ref{sec:model} we explain in details our model to realize Josephson junction with exciton polariton condensates.
The extended tight-binding model will be established in this case.
In Sec. \ref{sec:quantum_phase} we {\red \dots }.
In the conclusion, we summarize the results obtained.

\section{THE MODEL OD EXCITON POLARITON JOSEPHSON JUNCTION \label{sec:model}}

We start by describing the problem of equilibrium exciton polariton condensate trapping in symmetric one dimensional double-well potential $U(x) = h ((x/x_0)^2 - 1)^2$; $2x_0$ defines distance between potential minima, $h$ depth of the potential.
% 
%\begin{figure}[ht]
%\center{\includegraphics[width=0.5\linewidth]{pic/potential_sym_asym.png}}
%\caption{
%Confined double-well potential for exciton polariton coupled condensate channels with symmetric $(\Phi_+(x))$ and asymmetric $(\Phi_-(x))$ wavefunctions obtained numerically by solving stationary GPE (\ref{eq:stationary}).
%The dimensionless parameters for the calculations are: $h = 10$, $x_0 = 1$, $\mu_+ = 5.7355$, $\mu_- = 6.5653$.
%\label{pic:potential_sym_asym}

%\end{figure}
%
The Hamiltonian for the system of weakly interacting exciton polaritons in an external $U(x)$ can be written in the second quantized form as:
%
\begin{equation}
\hat{H} = \int \Big\{ -\dfrac{\hbar^2}{2m} \hat{\psi}^\dag \dfrac{d^2 \hat{\psi}}{dx^2} + \hat{\psi}^\dag \hat{\psi} U(x) + \dfrac{g}{2} \hat{\psi}^{\dag 2} \hat{\psi}^2  \Big\} dx,
\label{eq:gpe_hamiltonian}
\end{equation}
%
where $\hat{\psi} \equiv \hat{\psi}(x, t)$ is a polariton field operator that annihilate a particle at the position $x$ and time $t$, $g$ is two-body (polariton-polariton) scattering length.
In the paper we use so-called two-mode representation of $\hat{\psi}$ - operator supposing
%
\begin{equation}
\hat{\psi} = \hat{\psi}_1(t) \Phi_1(x) + \hat{\psi}_2(t) \Phi_2(x),
\label{eq:two_modes}
\end{equation}
%
where $\hat{\psi}_{1,2}$ are time dependent operators characterizing condensate at the left and at the right wells respectively; $\Phi_{1,2}(x)$ are condensate wavefunctions which are responsible for it spatial distribution.
It's instructive to recast $\Phi_{1,2}(x)$ as
%
\begin{equation}
\Phi_{1,2} = \dfrac{\Phi_+ \pm \Phi_-}{\sqrt{2}}
\label{eq:basic_modes}
\end{equation}
%
where $\Phi_{\pm}(x) = \pm \Phi_{\pm}(-x)$ are symmetric ($\Phi_+$) and antisymmetric ($\Phi_-$) real wavefunctions obeying stationary Gross-Pitaevskii equations $\mu_{\pm} \Phi_{\pm} = -\dfrac{\hbar^2}{2m} \dfrac{d^2 \Phi_{\pm}}{dx^2} + U(x) \Phi_{\pm} + g \Phi_{\pm}^3$.
%
%\begin{equation}
%\mu_{\pm} \Phi_{\pm} = -\dfrac{\hbar^2}{2m} \dfrac{d^2 \Phi_{\pm}}{dx^2} + U(x) \Phi_{\pm} + g \Phi_{\pm}^3.
%\label{eq:stationary}
%\end{equation}
%
and normalization condition $\int \Phi_{\pm}^2 dx = 1$.
Here $\mu_{\pm}$ are chemical potentials.
As a result an operator $\hat{\psi_1}^\dag\hat{\psi_1} + \hat{\psi_2}^\dag\hat{\psi_2} = N$ characterizes total number of particles (here and thereafter we omit ``hat'' design for simplicity).
Substituting (\ref{eq:basic_modes}), (\ref{eq:two_modes}) into (\ref{eq:gpe_hamiltonian}) we arrive to
% 
\begin{equation}
\begin{array}{lcl}
\hat{H} & = & \dfrac{A}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_1^2 + \hat{\psi}_2^{\dag 2} \hat{\psi}_2^2) - \dfrac{G}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 + \hat{\psi}_1 \hat{\psi}_2^\dag) \\ [8pt]
& & -\dfrac{\Gamma}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_1 \hat{\psi}_2 + \hat{\psi}_1^\dag \hat{\psi}_1^2 \hat{\psi}_2^\dag + \hat{\psi}_1^\dag \hat{\psi}_2^\dag \hat{\psi}_2^2 + \hat{\psi}_1 \hat{\psi}_2^{\dag 2} \hat{\psi}_2) \\ [8pt]
& & +\dfrac{C}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_2^2 + 4 \hat{\psi}_1^\dag \hat{\psi}_1 \hat{\psi}_2^\dag \hat{\psi}_2 + \hat{\psi}_1^2 \hat{\psi}_2^{\dag 2}),
\end{array}
\label{eq:hamiltonian}
\end{equation}
%
where we made some denotations $\gamma_{ij} = g \int \Phi_i^2 \Phi_j^2 dx, i,j \in \{+,-\}$;
$A = \frac{1}{4} (\gamma_{++} + \gamma_{--} + 6 \gamma_{+-})$; $\Gamma = \frac{1}{2} (\gamma_{++} - \gamma_{--})$; $C = \frac{1}{4} (\gamma_{++} + \gamma_{--} - 2\gamma_{+-})$; $G = \mu_- - \mu_+ - 2\Gamma$.

Hamiltonian (\ref{eq:hamiltonian}) represents extended model for internal Josephson junction effect with exciton polariton BEC trapped in double-well potential.
The ``additional'' terms with $\Gamma$ and $C$ in (\ref{eq:hamiltonian}) characterize density induced and pair tunneling, respectively.
Generally, our model implies existence of quantum tunneling of the particles between wells with effective (nonlinear) tunneling rate $G_{eff} = \dfrac{1}{2}(G+\Gamma N - C\hat{\psi}_1^\dag\hat{\psi}_2)$.
The contribution from relevant coefficients on can be estimated from variational ansatz
%
\begin{equation}
\Phi_{\pm} = A_{\pm} \Big[ \exp \Big( -\dfrac{(x - x_0)^2}{2 a^2} \Big) \pm \exp \Big( -\dfrac{(x + x_0)^2}{2 a^2} \Big) \Big]
\label{eq:two_modes_eq}
\end{equation}
%
where $a$ is width of condensate wavefunctions at each wells (we suppose them equal to each other), defined from normalization conditions.
In Fig. \ref{pic:gamma_pm_vs_g} the behavior of normalized coefficients $\dfrac{\gamma_{ij}}{g}$ $i,j \in \{+,-\}$ as a function of half of intra-well distance $x_0$ is shown.
It is important that at differences between them rapidly vanishes with increasing of distance $x_0$.
We can suppose $\Gamma = C = 0$ for $x_0 >> a$.
This limit exactly corresponds to familiar problem of two weakly linked condensates at zero temperature, see e.g. \cite{Aleiner, Shelykh_2008, Borgh_2010, Raghavan}.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.5\linewidth]{pic/ypm_x0.eps}}
\caption{Dependence of $\gamma_{ij} / g$ on $x_0 / a$ \label{pic:gamma_pm_vs_g}}
\end{figure}
%

@@

To elucidate purely quantum phase features of our extended Josephson model it is fruitful to recast Hamiltonian (\ref{eq:hamiltonian}) through the pseudo-spin operators defined as
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & \dfrac{1}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 + \hat{\psi}_1 \hat{\psi}_2^\dag) = s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & \dfrac{i}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 - \hat{\psi}_1 \hat{\psi}_2^\dag) = s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & \dfrac{1}{2} (\hat{\psi}_2^\dag \hat{\psi}_2 + \hat{\psi}_1^\dag \hat{\psi}_1) = -i\dfrac{d}{d \phi},
\end{align}
\label{eq:pseudo_spin}
\end{subequations}
%
where  $\phi$ is the phase variable.
The operators determined in Eq. (\ref{eq:pseudo_spin}) obey familiar $SU(2)$ algebra commutation relations $[S_i, S_j] = i\epsilon_{ijk}S_k,  i,j,k=x,y,z$.

>>>>

\author{M. E. Lebedev}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{D. A. Dolinina}
\affiliation{ITMO University, St. Petersburg 197101, Russia}

\author{S. M. Arakelian}
\affiliation{Vladimir State University named after A. G. and N. G. Stoletovs, Gorkii Street 87, Vladimir, Russia}

\author{Tien-Chang Lu}
\affiliation{Department of Photonics, National Chiao Tung University, Hsinchu 300, Taiwan}

\author{A. V. Kavokin}
\affiliation{Spin Optics Laboratory, St. Petersburg State University, Ul’anovskaya, Peterhof, St. Petersburg 198504, Russia}
\affiliation{School of Physics and Astronomy, University of Southampton, SO17 1BJ Southampton, United Kingdom}

\author{A. P. Alodjants}
\email{alexander\_ap@list.ru}
\affiliation{ITMO University, St. Petersburg 197101, Russia}
\affiliation{Vladimir State University named after A. G. and N. G. Stoletovs, Gorkii Street 87, Vladimir, Russia}

@@

\begin{abstract}
We consider the problem of achievement of quantum regime for  relative phase of trapped in W -- potential low branch exciton polariton condensates taken at finite temperatures and finite polariton lifetime. Non-standard Bose-Hubbard model of  exciton polariton Josephson junction (EPJJ) that produces complex potential energy landscape (PEL) consisting from set of barriers and wells for the phase is proposed. We have shown that classical (thermal) activation and quantum tunneling for the phase  represent two regimes depending on temperature, polariton condensate density and distance between the junctions. The transition between this two regimes   exhibits universal features of first or second order phase transition (PT) depending on the PEL for relative phase. In the presence of dissipation quantum system exhibits non-equilibrium PT from quantum (where tunneling is possible) to the classical regime where tunneling is escaped. The obtained results are discussed in details in the framework of quantum annealing problem with exciton polariton condensates.

\end{abstract}

@@

For last decade investigations of exciton polariton Bose-Einstein condensates (BEC) in various type of semiconductor microstructures represent important mainstream of current photonic and semiconductor technologies, see e.g. \cite{Sanvitto,Guillet}.
The microcavity exciton polaritons are bosonic quasiparticles representing  admixtures of the quantized cavity mode and quantum wells excitons.
Practically, semiconductor microcavities are promising for various optoelectronic applications where quantum  matter-field interface plays crucial role.
In particular, it is worth to mention demonstrations of polariton laser with electrical pump \cite{Bhattacharya,Schneider}, amplifiers \cite{Niemietz}, switches \cite{Amo_2010}, transistors \cite{Ballarini}, polariton circuits and optical logic elements \cite{Sturm,Liew}.
Although exciton polariton condensates exhibits Bose-Einstein statistics under the phase transition (PT) condition and macroscopic occupation of the ground state at certain pumping rate that is less than for convenient lasers, they are not in true thermodynamic equilibrium state \cite{Byrnes_2014,Sun}. Non-equilibrium features of exciton polariton condensates play important role in various manifestations of their collective (many body) states such as superfluidity \cite{Carusotto_2013,Amo_2009}, quantized vortices \cite{Lagoudakis_2008,Lagoudakis_2009}, soliton formation \cite{Sich}, Josephson oscillations and macroscopic self-trapping \cite{Abbarchi,Lagoudakis_2010}.

Recently Yongbao Sun et al,  in Ref. \cite{Snoke_2017} reports observation of quasi-equilibrium low branch exciton polariton condensate in high Q-factor microcavity possessing very high photon lifetime, that is $135ps$. Notably, exciton polaritons with long lifetime are promising candidates for current quantum technologies, cf. \cite{Demirchyan}. However, since exciton polariton BEC's in present experiments are observed at high enough (up to the room) temperatures the quantum abilities of such a states should be examined even at full thermal equilibrium.

The problem of distinguishability of statistically classical and quantum regimes for exciton polariton condensates behavior is very actual for their practical purposes and especially for their possible applications in quantum information science, see e.g. \cite{Dominici}.
Fast switching properties (the typical switching time of a few picoseconds), relatively strong nonlinear response and flexibility to external optical and/or electrical pump, spin degrees of freedom made microcavity polaritons potentially very promising for quantum computation and quantum information processing objectives, see, e.g. \cite{Demirchyan,Pagel,Kyriienko,Solnyshkov_2015, Dominici}.

Especially it is important to stress here quantum annealing problem physically based on searching algorithm for the global minimum of a potential energy landscape (PEL) consisting from set of barriers and wells, see e.g.  \cite{Santoro, Das}. In purely classical  (thermal) regime physical system overcomes arbitrary barrier at finite temperature with the help of thermal activation process if the thermal energy is enough. Contrary, in quantum limit the same system undergoes quantum tunneling through the barrier. Obviously, the shape of the barrier plays essential role in this case, cf. \cite{Das, Lewenstein}. In order, as it is shown recently  in \cite{Lewenstein} annealing algorithm implies in general case combination of thermal annealing and quantum tunneling. The     
collective (bosonic) character of polariton condensation could be used in this case to acceleration of search physical algorithm \cite{Yan}.

In the paper we consider generic problem  of quantum-classical phase transition (PT) with exciton polariton condensates occurring in semiconductor microcavity in the presence of Josephson effect, cf. \cite{Chudnovsky_1997,Aleiner, Shelykh_2008, Borgh_2010}. To be more specific we examine simply two junction model.

In general this problem relates to dissipative tunneling problem or, tunneling at finite (non-zero) temperatures, cf. \cite{Caldeira, Larkin, Riseborough}.
Traditionally,  studies in this field are limited by consideration of superconductor devices \cite{Ankerhold} where macroscopic quantum tunneling phenomena plays essential role.
In particular, it is worth to mention here Schr\"odinger cat state formation \cite{Leggett} and design of quantum gates with superconductor qubits for quantum computing where macroscopic quantum coherence for the phase is important \cite{Makhlin}.
Recently the problem of quantum tunneling and quantum-classical PT problem have been applied to other two-level (spin) systems, like atomic BEC's \cite{Zhang}, and small ferromagnetic particles, \cite{Owerre}.
Strictly speaking main discussion is focused on investigation of quantum-classical escape rate transition that takes place in the presence of potential barrier at finite temperatures.

In the paper we propose \textit{extended model} for exciton polariton Josephson junctions (EPJJs) that relates to so-called non-standard Bose-Hubbard models (see \cite{Dutta} and references therein) when energy of polariton-polariton scattering contributes into the tunneling parameter matrix elements, cf. \cite{Aleiner, Shelykh_2008, Borgh_2010, Solnyshkov_2008, Sarchi}.
In particularly, we pay our attention to temperature dependent quantum critical phenomena occurring in the presence of macroscopic tunneling.
We aimed at elucidation of necessary criteria to obtain quantum phase regimes with trapped exciton polaritons suitable for quantum annealing purposes and creation of exciton polariton qubit gates.

This paper is arranged as follows.
In Sec. \ref{sec:model} we explain in details  extended model of EPJJ under the tight-binding approach. 
In Sec. \ref{sec:quantum_phase} we give description of quantum phase for our model and discuss various regimes for it. Section \ref{sec:quantum_classical} establishes main results obtained in the paper. In order, we consider quantum-classical PT problem for various PEL's for quantum phase in the framework of our model at finite temperatures.  The influence of non-equilibrium effects with microcavity exciton polaritons possessing finite lifetime is discussed in \ref{sec:non-equilibrium}.      
In Conclusion we summarize the results obtained.

\section{THE MODEL OF EPJJ \label{sec:model}}

We start by describing the problem of equilibrium exciton polariton condensate trapping in symmetric one dimensional double-well potential $U(x) = h ((x/x_0)^2 - 1)^2$; $2x_0$ defines distance between potential minima, $h$ is depth of the potential.
% 
%\begin{figure}[ht]
%\center{\includegraphics[width=0.5\linewidth]{pic/potential_sym_asym.png}}
%\caption{
%Confined double-well potential for exciton polariton coupled condensate channels with symmetric $(\Phi_+(x))$ and asymmetric $(\Phi_-(x))$ wavefunctions obtained numerically by solving stationary GPE (\ref{eq:stationary}).
%The dimensionless parameters for the calculations are: $h = 10$, $x_0 = 1$, $\mu_+ = 5.7355$, $\mu_- = 6.5653$.
%\label{pic:potential_sym_asym}
%\end{figure}
%
The Hamiltonian for the system of weakly interacting exciton polaritons trapped in an external potential $U(x)$ can be written in the second quantized form as:
%
\begin{equation}
\hat{H} = \int \hat{\psi}^\dag\Big\{ -\dfrac{\hbar^2}{2m_{pol}}  \dfrac{d^2 }{dx^2} + U(x) + \dfrac{g}{2} \hat{\psi}^{\dag} \hat{\psi}  \Big\}\hat{\psi} dx,
\label{eq:gpe_hamiltonian}
\end{equation}
%
where $\hat{\psi} \equiv \hat{\psi}(x, t)$ is a polariton field operator that annihilate a particle at the position $x$ and time $t$, $g$ describes two-body (polariton-polariton) scattering length, $m_{pol}$ is low branch polariton effective mass.
In the paper we use so-called two-mode representation of $\hat{\psi}$ - operator supposing
%
\begin{equation}
\hat{\psi} = \hat{\psi}_1(t) \Phi_1(x) + \hat{\psi}_2(t) \Phi_2(x),
\label{eq:two_modes}
\end{equation}
%
where $\hat{\psi}_{1,2}$ are time dependent operators characterizing condensate at the left and at the right wells respectively; $\Phi_{1,2}(x)$ are condensate wavefunctions which are responsible for it spatial distribution.
It's instructive to represent  $\Phi_{1,2}(x)$ as
%
\begin{equation}
\Phi_{1,2} = \dfrac{\Phi_+ \pm \Phi_-}{\sqrt{2}}
\label{eq:basic_modes}
\end{equation}
%
where $\Phi_{\pm}(x) = \pm \Phi_{\pm}(-x)$ are symmetric ($\Phi_+$) and antisymmetric ($\Phi_-$) real wavefunctions obeying stationary Gross-Pitaevskii (GP) equations $\mu_{\pm} \Phi_{\pm} = -\dfrac{\hbar^2}{2m_{pol}} \dfrac{d^2 \Phi_{\pm}}{dx^2} + U(x) \Phi_{\pm} + g \Phi_{\pm}^3$
%
%\begin{equation}
%\mu_{\pm} \Phi_{\pm} = -\dfrac{\hbar^2}{2m} \dfrac{d^2 \Phi_{\pm}}{dx^2} + U(x) \Phi_{\pm} + g \Phi_{\pm}^3.
%\label{eq:stationary}
%\end{equation}
%
and normalization condition $\int \Phi_{\pm}^2 dx = 1$; $\mu_{\pm}$ is chemical potential.
As a result an operator $N=\hat{\psi_1}^\dag\hat{\psi_1} + \hat{\psi_2}^\dag\hat{\psi_2}$ characterizes total number of particles.
Substituting (\ref{eq:basic_modes}), (\ref{eq:two_modes}) into (\ref{eq:gpe_hamiltonian}) we arrive to
% 
\begin{equation}
\begin{array}{lcl}
\hat{H} & = & \dfrac{A}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_1^2 + \hat{\psi}_2^{\dag 2} \hat{\psi}_2^2) - \dfrac{G}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 + \hat{\psi}_1 \hat{\psi}_2^\dag) \\ [8pt]
& & -\dfrac{\Gamma}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_1 \hat{\psi}_2 + \hat{\psi}_1^\dag \hat{\psi}_1^2 \hat{\psi}_2^\dag + \hat{\psi}_1^\dag \hat{\psi}_2^\dag \hat{\psi}_2^2 + \hat{\psi}_1 \hat{\psi}_2^{\dag 2} \hat{\psi}_2) \\ [8pt]
& & +\dfrac{C}{2} (\hat{\psi}_1^{\dag 2} \hat{\psi}_2^2 + 4 \hat{\psi}_1^\dag \hat{\psi}_1 \hat{\psi}_2^\dag \hat{\psi}_2 + \hat{\psi}_1^2 \hat{\psi}_2^{\dag 2}),
\end{array}
\label{eq:hamiltonian}
\end{equation}
%
where we made denotations $\gamma_{ij} = g \int \Phi_i^2 \Phi_j^2 dx, i,j \in \{+,-\}$,
$A = \frac{1}{4} (\gamma_{++} + \gamma_{--} + 6 \gamma_{+-})$, $\Gamma = \frac{1}{2} (\gamma_{--} - \gamma_{++})$, $C = \frac{1}{4} (\gamma_{++} + \gamma_{--} - 2\gamma_{+-})$, $G = \mu_- - \mu_+ - 2\Gamma$.

Hamiltonian (\ref{eq:hamiltonian}) represents extended model for internal Josephson junction effect with exciton polariton condensates trapped in double-well potential.
The ``additional'' terms with $\Gamma$ and $C$ in (\ref{eq:hamiltonian}) characterize density induced and pair tunneling, respectively.
Generally, our model implies existence of quantum tunneling of the particles between wells with effective (nonlinear) tunneling rate $G_{eff} = \dfrac{1}{2}(G+\Gamma N - C\hat{\psi}_1^\dag\hat{\psi}_2)$.
The contribution from relevant coefficients on can be estimated from variational ansatz
%
\begin{equation}
\Phi_{\pm} = A_{\pm} \Big[ \exp \Big( -\dfrac{(x - x_0)^2}{2 a^2} \Big) \pm \exp \Big( -\dfrac{(x + x_0)^2}{2 a^2} \Big) \Big],
\label{eq:two_modes_eq}
\end{equation}
%
where $a$ is width of condensate wavefunctions at each wells (we suppose them equal to each other), defined from normalization conditions.
In Fig. \ref{pic:gamma_pm_vs_g} the behavior of normalized coefficients $\dfrac{\gamma_{ij}}{g}$ $i,j \in \{+,-\}$ as a function of half of inter-well distance $x_0$ is shown.
It is important that differences between various $\gamma_{ij}$ rapidly vanish with increasing the inter-well distance $2x_0$.
We can suppose $\Gamma = C = 0$ for $x_0 >> a$.
This limit exactly corresponds to familiar problem of two weakly linked condensates at zero temperature, see e.g. \cite{Aleiner, Shelykh_2008, Borgh_2010, Raghavan}.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.5\linewidth]{pic/ypm_x0.eps}}
\caption{Dependences of normalized matrix elements $\gamma_{ij} / g$ on $x_0 / a$. \label{pic:gamma_pm_vs_g}}
\end{figure}
%

@@

To elucidate purely quantum phase features of our extended Josephson model it is fruitful to recast Hamiltonian (\ref{eq:hamiltonian}) through the pseudo-spin operators defined as
% 
\begin{subequations}
\begin{align}
\hat{S}_x = & \dfrac{1}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 + \hat{\psi}_1 \hat{\psi}_2^\dag) = s \cos{\phi} - \sin{\phi} \dfrac{d}{d \phi}, \\
\quad \hat{S}_y = & \dfrac{i}{2} (\hat{\psi}_1^\dag \hat{\psi}_2 - \hat{\psi}_1 \hat{\psi}_2^\dag) = s \sin{\phi} + \cos{\phi} \dfrac{d}{d \phi}, \\
\hat{S}_z = & \dfrac{1}{2} (\hat{\psi}_2^\dag \hat{\psi}_2 - \hat{\psi}_1^\dag \hat{\psi}_1) = -i\dfrac{d}{d \phi},
\end{align}
\label{eq:pseudo_spin}
\end{subequations}
where  $\phi$ is the phase variable.
The operators determined in Eq. (\ref{eq:pseudo_spin}) obey familiar $SU(2)$ algebra commutation relations $[S_i, S_j] = i\epsilon_{ijk}S_k,  i,j,k=x,y,z$.
After straitforward calculations from (\ref{eq:hamiltonian}) one can obtain
%
\begin{equation}
\begin{array}{lcl}
H & = & -(\alpha - \beta \sin^2 \phi) \dfrac{d^2}{d \phi^2} - (\beta(s-\dfrac{1}{2})\sin{2\phi} - B\sin{\phi}) \dfrac{d}{d \phi} \\
&& - Bs \cos \phi - \beta s^2 \sin^2 \phi - \beta s \cos^2 \phi.
\end{array}
\label{eq:hamiltinian_s}
\end{equation}
%
where $\alpha = A - C = 2\gamma_{\pm} > 0$, $\beta = 2C > 0$ and $B = \Gamma N + G > 0$.
Below we chose $x_0$ obeying to condition $\gamma_{++} \simeq \gamma_{--} = \gamma$ supposing $\Gamma = 0$ for simplicity, see Fig. \ref{pic:gamma_pm_vs_g}. We also assume macroscopically large particle number, i.e. the $s = N/2 >> 1$ is C - number. Using (\ref{eq:hamiltinian_s})  for Schrodinger equation we obtain   
%
\begin{equation}
\begin{array}{l}
(\alpha - \beta \sin^2 \phi) \dfrac{d^2 \Phi}{d \phi^2} + (\beta s\sin{2\phi}-B\sin{\phi})) \dfrac{d \Phi}{d \phi} \\
+ (E + B s \cos \phi + \beta s^2 \sin^2 \phi) \Phi = 0,
\end{array}
\label{eq:schrodinger}
\end{equation}
%
where $\Phi \equiv \Phi(\phi)$ is unknown $2\pi$-periodic wavefunction that characterizes quantum phase properties, cf. \cite{Anglin}.
%ng that Eqs. (\ref{eq:hamiltinian_s}), (\ref{eq:schrodinger}) can be obtained by applying so-called phase-state representation, {\red cf. [  ]}.
%
%\begin{equation}
%\ket{\psi} = \dfrac{1}{2 \pi} \int_{-\infty}^{+\infty} d \phi \Phi(\phi) \ket{\phi}
%\end{equation}
%
%where
%
%\begin{equation}
%\ket{\phi} = \dfrac{1}{\sqrt{N!}} \Big[ e^{i\phi/2} \psi_1^\dagger + e^{-i \phi/2} \psi_2^\dag \Big]^N \ket{0}_1\ket{0}
%\end{equation}
%
%is macroscopic coherent spin-state (macroscopic qubit state) that plays important role in current quantum information protocols operating with $N$-particle condensates, cf. \cite{Byrnes_2012}.
It is possible to eliminate the term with first derivative in Eq. (\ref{eq:schrodinger}) applying the transformation
%
\begin{equation}
\Phi(\phi) = \Psi(z) \exp \Big( s \ln \dn{z} - \dfrac{\Lambda s}{2 \sqrt{\lambda (1 - \lambda)}} \arctan \Big( \sqrt{\dfrac{\lambda}{1 - \lambda}} \cn{z} \Big) \Big),
\label{eq:transformation_of_phi}
\end{equation}
%
where $z = \int \limits_0^\phi \dfrac{d \xi}{\sqrt{1 - \lambda \sin^2 \xi}} = F(\phi, \lambda)$ is new phase variable; $F(\phi, \lambda)$ is incomplete elliptic integral of the first kind.
In (\ref{eq:transformation_of_phi}) we have introduced dimension-less parameters $\Lambda = \dfrac{G}{\alpha s}$, $\lambda = \dfrac{\beta}{\alpha}$ playing crucial role for our model.
Inserting (\ref{eq:transformation_of_phi}) into (\ref{eq:schrodinger}) we arrive to Schr\"odinger-like equation
% 
\begin{equation}
\alpha \frac{d^2\Psi}{dz^2} + (E - V(z))\Psi = 0
\label{eq:schrodinger_usual}
\end{equation}
%
characterizing quantum particle with the ``mass''  $m = \dfrac{\hbar^2}{2 \alpha}$ and energy $E$, that moves in the potential $V(z) = \alpha s^2 V_0(z)$ with
%
\begin{equation}
V_0(z) = \frac{ (\frac{1}{4} \Lambda^2 - \lambda (1 - \lambda)) \sn^2{z} - \Lambda \cn{z}}{\dn^2{z}}.
\label{eq:potential}
\end{equation}
%

The dependences of trapping potential $V_0(z)$ on phase variable $z$ are shown in Fig. \ref{phase_potential}. The period of the functions is  $4K(\lambda)$.
The dependence of the $\lambda$ parameter on normalized half of inter-well distance $\dfrac{x_0}{a}$ can be estimated from (\ref{eq:two_modes_eq}) and reads as $\lambda = 0.5 \Big( \exp \Big[ \dfrac{2 x_0^2}{a^2} \Big] - 1 \Big)^{-1}$. The results known for quantum phase mesoscopic Josephson junction model can be recovered from (\ref{eq:hamiltonian}), (\ref{eq:potential}) at $\lambda = 0$ and shown by red curves in  Fig.\ref{phase_potential}. This limit corresponds to  infinitely large inter-well distances, with $x_0 \to \infty$.
On the other hand, $\lambda$ -- parameter grows rapidly at $\dfrac{x_0}{a} \ll 1$.
Obviously, in this limit our EPJJ model that based on assumption of relatively weak coupling between the wells breaks down.

Below we are focusing on $\lambda$ -- parameters which belong to segment of $0 < \lambda < 1$ and might be obtained at moderate values of $\dfrac{x_0}{a}$ which are $\dfrac{x_0}{a} \geq 0.45$.
